---
title: "Tables with datagrid()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{datagrid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(toastui)
```


## Usage

Create a table from a `data.frame` with `datagrid()`:

```{r default-usage}
datagrid(rolling_stones_50)
```



## Column sizing

Default column width above is `"fit"`, to (try to) fit container width (depending on the number of columns), other alternatives are:

* `"guess`: to guess minimum width according to content and column names.
* `"auto"`: to set width accordingto content, excluding column names.
* `NULL`: to do nothing and use default behavior.

```{r colwidth-alternatives}
datagrid(rolling_stones_50, colwidths = "guess")
datagrid(rolling_stones_50, colwidths = "auto")
datagrid(rolling_stones_50, colwidths = NULL)
```

It's also possible to set width in pixel for columns with `grid_columns()`, here we set width for first two variables to 150px:

```{r}
datagrid(rolling_stones_50) %>% 
  grid_columns(vars = c("Number", "Year"), width = 150)
```


## Height

Default height of table created is 600px (in markdown and shiny), you can use a different value by using:

```{r height}
datagrid(rolling_stones_50, height = "300px")
```

To display all rows in table use `bodyHeight` argument:

```{r body-heigh}
datagrid(rolling_stones_50, bodyHeight = "auto")
```




An other solution to control row displayed is to use pagination.


## Pagination

Activate pagination by indicating the number of rows to display:

```{r}
datagrid(rolling_stones_50, pagination = 8)
```



## Sort

Sort is activated by default, you can desactivate like this:

```{r}
datagrid(rolling_stones_50, sortable = FALSE)
```



## Filter

You can allow filtering rows in the table with:

```{r}
datagrid(rolling_stones_50, filters = TRUE)
```


You can customize filter type with `grid_filters()`:

```{r}
datagrid(rolling_stones_50) %>% 
  grid_filters(
    vars = "Artist", 
    showApplyBtn = TRUE, 
    showClearBtn = TRUE, 
    type = "text"
  )
```


## Edit content

### Editor options

You can edit data in table and retrieve back results in Shiny, double click on any cell to see different options:

```{r}
edit <- data.frame(
  col_text = month.name,
  col_select = month.name,
  col_checkbox = month.abb,
  col_radio = month.name,
  col_month = format(as.Date("2020-01-01") + 0:11*31, format = "%Y-%m")
)
datagrid(edit, bodyHeight = "auto") %>% 
  grid_editor("col_text", type = "text") %>% 
  grid_editor("col_select", type = "select", choices = month.name) %>% 
  grid_editor("col_checkbox", type = "checkbox", choices = month.abb) %>% 
  grid_editor("col_radio", type = "radio", choices = month.name) %>% 
  grid_editor_date("col_month", type = "month", format = "yyyy-MM")
```

In Shiny, retrieve data edited with `input$<outputId>_data`.


### Validation

When grid is editable you can set some rules to validate what values user can enter:

```{r validation}
validate <- data.frame(
  col_text = c("a", "b", "a", NA, "c"),
  col_number = sample(1:10, 5),
  col_mail = c("victor@mail.com", "victor", NA, "victor@mail", "victor.fr")
)

datagrid(validate, bodyHeight = "auto") %>%
  grid_editor(
    "col_text", type = "text",
    validation = validateOpts(required = TRUE, unique = TRUE)
  ) %>%
  grid_editor(
    "col_number", type = "number",
    validation = validateOpts(min = 0, max = 5)
  ) %>%
  grid_editor(
    "col_mail", type = "text",
    validation = validateOpts(
      regExp = "^([a-zA-Z0-9_\\-\\.]+)@([a-zA-Z0-9_\\-\\.]+)\\.([a-zA-Z]{2,5})$"
    )
  )
```

In Shiny, retrieve validation errors with `input$<outputId>_validation`.


## Selection

In Shiny applications, you can select multiple rows with checkbox, single row with radio button or select cells by holding and selecting an area.

```{r selection}
# Checkbox
datagrid(rolling_stones_50) %>%
  grid_selection_row(
    inputId = "sel_check",
    type = "checkbox"
  )

# Radio
datagrid(rolling_stones_50) %>%
  grid_selection_row(
    inputId = "sel_radio",
    type = "radio"
  )
```



## Row & Column groups

Group columns under a common label:

```{r col-groups}
datagrid(ps3_games[order(ps3_games$Publisher), c(1, 4, 5:8)], colwidths = "guess") %>%
  grid_columns(
    vars = c("NA_Sales", "EU_Sales", "JP_Sales", "Other_Sales"),
    header = c("North America", "Europe", "Japan", "Other")
  ) %>%
  grid_complex_header(
    "Sales"= c("NA_Sales", "EU_Sales", "JP_Sales", "Other_Sales")
  )
```


Group rows with common values:

```{r row-groups}
datagrid(ps3_games[order(ps3_games$Publisher), c(1, 4, 5:8)], colwidths = "guess") %>%
  grid_row_merge(vars = "Publisher")
```



## Summary

Add a summary area to the bottom of the grid to display statistics computed on column:

```{r summary-area}
datagrid(ps3_games[, c(1, 5:8)], colwidths = "guess") %>% 
  grid_summary(
    column = c("NA_Sales", "EU_Sales", "JP_Sales", "Other_Sales"),
    stat = "sum",
    label = "Total: "
  )
```




## Styling grid

### Color bar

```{r colorbar}
data.frame(
  rn = rownames(mtcars),
  mpg = mtcars$mpg,
  disp = mtcars$disp
) %>% 
  datagrid(
    colnames = c("Automobile", "Miles/(US) gallon", "Displacement (cu.in.)")
  ) %>%
  grid_colorbar(
    column = "mpg",
    label_outside = TRUE,
    label_width = "25px"
  ) %>%
  grid_colorbar(
    column = "disp",
    bar_bg = "#BF616A"
  )
```


## Theme

It's possible to customize theme used to create tables, theme is set globally and will be applied to all table created after its declaration.
In a rmarkdown document (like this one), theme is unique for all tables, click buttons to switch between a custom theme and the default one, you will see that all tables are styled.

```{r theme, eval=FALSE}
set_grid_theme(
  row.even.background = "#ddebf7",
  cell.normal.border = "#9bc2e6",
  cell.normal.showVerticalBorder = TRUE,
  cell.normal.showHorizontalBorder = TRUE,
  cell.header.background = "#5b9bd5",
  cell.header.text = "#FFF",
  cell.selectedHeader.background = "#013ADF",
  cell.focused.border = "#013ADF"
)

datagrid(rolling_stones_50)
```

```{r, echo=FALSE}
library(htmltools)
tags$div(
  tags$button(
    class = "btn btn-default",
    id = "apply-theme-custom",
    "Apply custom theme"
  ),
  tags$button(
    class = "btn btn-default",
    id = "apply-theme-default",
    "Back to default theme"
  )
)
```

```{js, echo=FALSE}
document.querySelector('#apply-theme-custom').addEventListener('click', function() {
  const options = {"selection":{},"scrollbar":{},"outline":{},"frozenBorder":{},"area":{"header":{},"body":{},"summary":{}},"row":{"odd":{},"even":{"background":"#ddebf7"},"dummy":{},"hover":{}},"cell":{"normal":{"border":"#9bc2e6","showVerticalBorder":true,"showHorizontalBorder":true},"header":{"background":"#5b9bd5","text":"#FFF"},"rowHeader":{"border":"#9bc2e6","showVerticalBorder":true,"showHorizontalBorder":true},"summary":{},"selectedHeader":{"background":"#013ADF"},"selectedRowHeader":{},"focused":{"border":"#013ADF"},"focusedInactive":{},"required":{},"editable":{},"disabled":{},"invalid":{}}};
  tui.Grid.applyTheme('clean', options);
});
document.querySelector('#apply-theme-default').addEventListener('click', function() {
  tui.Grid.applyTheme('clean', {});
});
```

```{r, echo=FALSE}
datagrid(rolling_stones_50)
```

