---
title: "Calendars"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calendars}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(toastui)
```


## Usage

Create interactive calendars with:

```{r, eval=FALSE}
calendar()
```

You can display calendars in interactive R Markdown documents or Shiny applications:

* R Markdown: display calendars that you can navigate and click on schedules to display extra information.
* Shiny: display calendars where you can edit, update and drag-and-drop schedules.


## Monthly calendar

Create a monthly calendar:

* `cal_demo_data()` return a `data.frame` with schedules example
* `useNavigation = TRUE`: add navigation buttons to change the month displayed
* `cal_month_options()`: global options for calendar
* `cal_props()`: set properties for schedules groups (background and text color ...)

```{r}
calendar(cal_demo_data(), useNavigation = TRUE) %>%
  cal_month_options(
    startDayOfWeek  = 1, 
    narrowWeekend = TRUE
  ) %>% 
  cal_props(
    list(
      id = 1,
      name = "PERSO",
      color = "white",
      bgColor = "firebrick",
      borderColor = "firebrick"
    ),
    list(
      id = 2,
      name = "WORK",
      color = "white",
      bgColor = "forestgreen",
      borderColor = "forestgreen"
    )
  )
```



## Weekly calendars

```{r}
calendar(cal_demo_data("week"), view = "week") %>% 
  cal_week_options(
    startDayOfWeek = 1,
    workweek = TRUE
  ) %>% 
  cal_props(
    list(
      id = 1,
      name = "COURSES",
      color = "white",
      bgColor = "steelblue",
      borderColor = "steelblue"
    ),
    list(
      id = 2,
      name = "WORK",
      color = "white",
      bgColor = "forestgreen",
      borderColor = "forestgreen"
    ),
    list(
      id = 3,
      name = "WORK",
      color = "firebrick",
      bgColor = "#D8D8D8",
      borderColor = "firebrick"
    )
  )
```




